### 2025-03-15 quarto/lda.qmd

---
title: " LDA example"
author: "Ondrej Such"
date: "2025-03-11"
format: pdf
---

## Number of classes

First, let us read the data files
```{r read data files for mnist}
source("lda.R")
dfs <- read_wlws(800, "mnist", 2)
```

Now let us compute result of coupling methods for all triplets of classes
```{r triplets}
r <- lda_triples(dfs)
```

Now we display interesting cases

```{r display interesting cases}
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(knitr))
df <- r |> 
    pivot_wider(names_from = "method", values_from = "acc") |> 
    filter(wlw2 > normal + 0.15) |> 
    mutate(imp =  wlw2 - pmax(omit12, omit23, omit13)) |> 
    arrange(desc(imp))

kable(df, digits = 3)
```