### 2025-08-14 tab-exp4.tex

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Mon Jun 16 08:56:49 2025
\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrrr}
  \hline
method & dna & letter & mnist & satimage & segment & usps & waveform \\ 
  \hline
normal & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 \\ 
  radial & 0.022 & 0.039 & 0.020 & 0.014 & 0.009 & 0.014 & 0.002 \\ 
  wlw2 & 0.015 & 0.044 & 0.030 & 0.013 & 0.007 & 0.020 & 0.002 \\ 
   \hline
\end{tabular}
\caption{Comparison of the difference in accuracies between methods $\boldsymbol{c}_1$ and
             $\boldsymbol{c}_2$} 
\label{tab:exp4}
\end{table}