### 2025-08-14 tab-sep.tex

% latex table generated in R 4.5.1 by xtable 1.8-4 package
% Wed Aug 13 15:27:15 2025
\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
Dataset & $K$ & Percentage linearly separable \\ 
  \hline
dna &    3 & 100 \%    \\ 
  letter &   26 & 99 \%    \\ 
  mnist &   10 & 100 \%    \\ 
  satimage &    6 & 84 \%    \\ 
  segment &    7 & 89 \%    \\ 
  usps &   10 & 100 \%    \\ 
  waveform &    3 & 0 \%    \\ 
   \hline
\end{tabular}
\caption{The benchmark datasets with indication of the number of classes $K$ and the percentage of binary classification problems that are linearly separable} 
\label{tab:sep}
\end{table}