### 2025-08-14 tab-step2.tex

% latex table generated in R 4.5.1 by xtable 1.8-4 package
% Thu Aug 14 09:38:32 2025
\begin{table}[ht]
\centering
\begin{tabular}{lr}
  \hline
Outcome & Percentage \\ 
  \hline
 $\boldsymbol{e}_3$ at least as accurate as $\boldsymbol{wlw}$ & 24 \%   \\ 
  one of $\boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$ at least as accurate as $\boldsymbol{wlw}$ & 53 \%   \\ 
  $\boldsymbol{wlw}$ more accurate than $\boldsymbol{e}_3, \boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$ & 39 \%   \\ 
   \hline
\end{tabular}
\caption{Comparison of Wu--Lin--Weng's method $\boldsymbol{wlw}$ and  parameterless coupling methods on three class subsets. } 
\label{tab:step2}
\end{table}